---
title: "Lab 4 Homework"
author: "Alejandra Ramirez"
date: "2/1/20"
output:
  html_document: 
    keep_md: yes
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove any `#` for included code chunks to run.  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

For this assignment we are going to work with a large dataset from the [United Nations Food and Agriculture Organization](http://www.fao.org/about/en/) on world fisheries. First, load the data.  

1. Load the data `FAO_1950to2012_111914.csv` as a new object titled `fisheries`.

```{r}
fisheries <- readr::read_csv("data/FAO_1950to2012_111914.csv")
```



2. What are the names of the columns? Do you see any potential problems with the column names?

```{r}
names(fisheries)
```
####yes, years should be part of the data

3. What is the structure of the data? Show the classes of each variable.

```{r}
str(fisheries)
```

```{r}
glimpse(fisheries)
```



4. Think about the classes. Will any present problems? Make any changes you think are necessary below.

####The years are under characters when they should be numeric
```{r}
fisheries<-
  fisheries%>%
  mutate_at(vars(starts_with("19")), as.numeric) %>% 
  mutate_at(vars(starts_with("2")), as.numeric) %>% 
  mutate_if(is.character, as.factor)

glimpse(fisheries)

#From your glimpse output it looks like this didn't work? (years are still dbl class) A more direct option would be
#fisheries$Country <- as.factor(fisheries$Country)
#fisheries$`ISSCAAP group#` <- as.factor(fisheries$`ISSCAAP group#`)
#etc...
#And we'll deal with the years later
```

5. How many countries are represented in the data? Provide a count.

```{r}
nlevels(fisheries$Country)
```




6. What are the names of the countries?

```{r}
levels(fisheries$Country)
```



7. Use `rename()` to rename the columns and make them easier to use. The new column names should be:  
+ country
+ commname
+ ASFIS_sciname
+ ASFIS_spcode
+ ISSCAAP_spgroup
+ ISSCAAP_spgroupname
+ FAO_area
+ unit


```{r}
colnames(fisheries)
```


##Formatting the code this way makes it more readable, and allows you to see inconsistancies in your punctuation

```{r}
Fisheries <-
  fisheries%>%
  dplyr::rename(country='Country', 
                commname='Common name', 
                ASFIS_sciname='ASFIS species name', 
                ASFIS_spcode='ASFIS species#', 
                ISSCAAP_spgroupname='ISSCAAP taxonomic group', 
                ISSCAAP_spgroup='ISSCAAP group#',  
                FAO_area='FAO major fishing area', 
                unit='Measure')
names(Fisheries)
#note that having multiple datasets whose name only varies by capitalization could get confusing
```


```{r}
colnames(Fisheries)
```

 



8. Are these data tidy? Why or why not, and, how do you know? Use the appropriate pivot function to tidy the data. Remove the NA's. Put the tidy data frame into a new object `fisheries_tidy`.  
##Make sure to answer the questions in your markdown file

```{r}
fisheries_tidy <-
  Fisheries %>%
  pivot_longer(-c(country, commname, ASFIS_sciname, ASFIS_spcode, ISSCAAP_spgroupname, ISSCAAP_spgroup, FAO_area, unit) ,
  names_to = "years", 
 values_to = "catch", values_drop_na = TRUE)
fisheries_tidy

```


9. Refocus the data only to include country, ISSCAAP_spgroupname, ASFIS_spcode, ASFIS_sciname, year, and catch.


```{r}
fisheries_tidy2 <-
  fisheries_tidy%>%
  select(country, ISSCAAP_spgroupname, ASFIS_spcode, ASFIS_sciname, years, catch)
fisheries_tidy2
```



10. Re-check the classes of `fisheries_tidy2`. Notice that "catch" is shown as a character! This is a problem because we will want to treat it as a numeric. How will you deal with this?

```{r}
glimpse(fisheries_tidy2)
```

#catch isn't a character
##But it still isn't a numeric so it still needs to be changed

```{r}
fisheries_tidy2$catch <- as.numeric(fisheries_tidy2$catch)
```


11. Based on the ASFIS_spcode, how many distinct taxa were caught as part of these data?

```{r}
nlevels(fisheries_tidy2$ASFIS_spcode)
```



12. Which country had the largest overall catch in the year 2000?
##The top line of this output is only for herrings and sardines. You needs to gruop by country and sum up the total catch
```{r}
fisheries_tidy2%>%
  filter(years==2000) %>%
  #group_by(country) %>% 
  #summarize(catch_total=sum(catch, na.rm=T)) %>% 
  arrange(desc(catch))
 
```

###Peru had the largest overall catch

13. Which country caught the most sardines (_Sardina_) between the years 1990-2000?
##Same as above, this code won't give you total catches. Also we can grab a range of years a little more efficiently
```{r}
fisheries_tidy2%>%
  filter(years%in% c(1990,1991,1992,1992,1994,1995,1996,1997,1998,1999,2000), str_detect(ASFIS_sciname,"Sardina"))%>%
  arrange(desc(catch))
  #filter(str_detect(ASFIS_sciname, "Sardina")) %>% 
  #filter(year>=1990 & year<=2000) %>% 
  #group_by(country) %>% 
  #summarize(catch_total=sum(catch, na.rm=T)) %>% 
  #arrange(desc(catch_total))
  
```

###Morocco had largest catch of sardines


14. Which five countries caught the most cephalopods between 2008-2012?
##Same suggestion for the years as above. 
##By only looking at the sciname you are actually only picking lines that couldn't be identified past class Cephalopoda. Pulling the ISSCAAP_spgroupname "Squids, cuttlefishes, octopuses" would get you the answer we're after 14.
##Your summarise has overwritten your catch variable, it should get a new name so that you don't lose data
```{r}
fisheries_tidy2 %>%
  filter(years %in% c(2008,2009,2010,2011,2012), str_detect(ASFIS_sciname, "Cephalopoda"))%>%
  group_by(country)%>%
  summarise(catch = sum(catch))%>%
  arrange(desc(catch))
  #filter(str_detect(ISSCAAP_spgroupname, "Squids")) %>% 
  #filter(year>=2008 & year<=2012) %>% 
  #group_by(country) %>% 
  #summarize(catch_total=sum(catch, na.rm=T)) %>% 
  #arrange(desc(catch_total))



```

###Top 5 are India, China, Cambodia, Madagascar, and Mozambique


15. Given the top five countries from question 12 above, which species was the highest catch total? The lowest?
##This question is still referring to cephalopods, so you could use the code above and then group by sciname like you have here
##Also note that your "catch" variable doesn't exist anymore because you overwrote it previously
```{r}
fisheries_tidy2%>%
  filter(country %in% c("India", "China", "Cambodia", "Madagascar", "Mozambique"))%>%
  group_by(ASFIS_sciname)%>%
  summarise(catch=sum(catch))%>%
  arrange(desc(catch))
              
  

           
```

####Osteichthyes was the highest catch total and Tetrapturus pfluegeri was the lowest catch

16. In some cases, the taxonomy of the fish being caught is unclear. Make a new data frame called `coastal_fish` that shows the taxonomic composition of "Miscellaneous coastal fishes" within the ISSCAAP_spgroupname column.
##Just a thought, if you group by sciname here, your output will look a little cleaner
```{r}
coastal_fish <- 
  fisheries_tidy2%>%
  filter(str_detect(ISSCAAP_spgroupname, "Miscellaneous coastal fishes"))

coastal_fish
```


17. Use the data to do at least one exploratory analysis of your choice. What can you learn?
##"What can you learn?" What is this code showing us?
```{r}
coastal_fish%>%
  filter(years>=2000)%>%
  filter(str_detect(ASFIS_sciname, "Boops boops"))%>%
  arrange(catch)
```




## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   






