---
title: "Lab 7 Homework"
author: "Alejandra Ramirez"
date: "2/25/2020"
output:
  html_document: 
    keep_md: yes
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for any included code chunks to run.  

## Libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(shiny)
library(shinydashboard)
```

## Data
The data for this assignment come from the [University of California Information Center](https://www.universityofcalifornia.edu/infocenter). Admissions data were collected for the years 2010-2019 for each UC campus. Admissions are broken down into three categories: applications, admits, and enrollees. The number of individuals in each category are presented by demographic.  
```{r}
UC_admit <- readr::read_csv("data/UC_admit.csv")
```

**1. Use the function(s) of your choice to get an idea of the overall structure of the data frame, including its dimensions, column names, variable classes, etc. As part of this, determine if there are NA's and how they are treated.**  

```{r}
UC_admit
```


```{r}
colnames(UC_admit)
glimpse(UC_admit)
```

```{r}
UC_admit%>%
  na_if("  ")%>%
  na_if("-999")%>%
  anyNA()
```

```{r}
UC_admit%>%
  summarise(total_nas = sum(is.na(UC_admit)))
```


**2. The president of UC has asked you to build a shiny app that shows admissions by ethnicity across all UC campuses. Your app should allow users to explore year, campus, and admit category as interactive variables. Use shiny dashboard and try to incorporate the aesthetics you have learned in ggplot to make the app neat and clean.**

```{r}
UC_admit%>%
  group_by(Campus)%>%
  count()
```


##Note you've put in a slider for adjusting point size but your graph doesn't have points.
##Also year is being treater as a numeric instead of a factor
```{r}
ui <- dashboardPage(
  dashboardHeader(title = "Plot UC Admissions App"),
  dashboardSidebar(),
  dashboardBody(
  fluidRow(
  box(title = "Plot Options", width = 3,
   selectInput("x", "Select X Variable", choices = c("Academic_Yr", "Campus", "Category"), 
              selected = "Academic_Yr"),
  sliderInput("pointsize", "Select the Point Size", min = 1, max = 5, value = 2, step = 0.5)
  ), # close the first box
  box(title = "Plot of UC Admissions Data", width = 7,
  plotOutput("plot", width = "600px", height = "500px")
  ) # close the second box
  ) # close the row
  ) # close the dashboard body
) # close the ui

server <- function(input, output, session) {
  
    # the code to make the plot of iris data grouped by species
  output$plot <- renderPlot({
    ggplot(UC_admit, aes_string(x = input$x , y = "FilteredCountFR" , fill = "Ethnicity" )) + geom_bar(stat= "identity", position = "dodge")
  })
  
  # stop the app when we close it
  session$onSessionEnded(stopApp)
  }

  
  shinyApp(ui, server)
```
  
**3. Make alternate version of your app above by tracking enrollment at a campus over all of the represented years while allowing users to interact with campus, category, and ethnicity.**
##This app should allow you to look at individual campuses
##Also the code for the slider has been left in here as well.
```{r}
ui <- dashboardPage(
  dashboardHeader(title = "Plot UC Admissions App Alternative"),
  dashboardSidebar(),
  dashboardBody(
  fluidRow(
  box(title = "Plot Options", width = 3,
   selectInput("x", "Select X Variable", choices = c("Ethnicity", "Campus", "Category"), 
              selected = "Ethnicity"),
  sliderInput("pointsize", "Select the Point Size", min = 1, max = 5, value = 2, step = 0.5)
  ), # close the first box
  box(title = "Plot of UC Admissions Data", width = 7,
  plotOutput("plot", width = "600px", height = "500px")
  ) # close the second box
  ) # close the row
  ) # close the dashboard body
) # close the ui

server <- function(input, output, session) {
  
    # the code to make the plot of iris data grouped by species
  output$plot <- renderPlot({
    ggplot(UC_admit, aes_string(x = "Academic_Yr"  , y = "FilteredCountFR"  , fill = input$x )) + geom_bar(stat= "identity", position = "dodge")
  })
  
  # stop the app when we close it
  session$onSessionEnded(stopApp)
  }

  
  shinyApp(ui, server)
```
##There are no independent apps in the folder. Though the do run directly from the code.



## Push your final code to GitHub!


Please be sure that you check the `keep md` file in the knit preferences. 